Shedding Light on the ``Dark Side'' of Phylogenetic Comparative Methods

### 2015-12-23 Manuscript/dark2.tex
Fixed BOH's comments

\parindent = 1.5em
\addtolength{\parskip}{.3em}
% Abstract 350 words max
\section{Abstract}
\begin{enumerate}
  \item Phylogenetic comparative methods are becoming increasingly popular for investigating evolutionary patterns and processes.
  However, these methods are not infallible - they suffer from biases and make assumptions like all other statistical methods.
  \item Unfortunately, although these limitations are generally well-known in the phylogenetic comparative methods community, they are often inadequately assessed in empirical studies leading to misinterpreted results and poor model fits.
  Here we explore reasons for the communication gap dividing those developing new methods and those using them. 
  \item We suggest that some important pieces of information are missing from the literature, and that others are difficult to extract from long, technical papers.
  We also highlight problems with users jumping straight into software implementations of methods (e.g. in R) that may lack documentation on biases and assumptions that are mentioned in the original papers.
  \item To help solve these problems we make a number of suggestions including providing
  blog posts or videos to explain new methods in less technical terms, encouraging reproducibility and code sharing, making wiki-style pages summarising the literature on popular methods, more careful consideration and testing of whether a method is appropriate for a given question/dataset, increased collaboration, and a shift from publishing purely novel methods to publishing improvements to existing methods and ways of detecting biases or testing model fit.
  Many of these points are applicable across methods in ecology and evolution, not just phylogenetic comparative methods.
\end{enumerate}

@@

We suggest that one cause of this problem is that although researchers developing and implementing new methods are aware of the limitations of their methods and the assumptions that underly them, this information is not always being effectively transferred to end-users.
Additionally, the tools and approaches used to fit models are often far more user-friendly and better documented than the methods used to assess whether that model fit is reasonable. 
Clearly more effort is needed to bridge the widening gap between those developing methods and end-users. 
Here we explore the causes of this communication gap and suggest some potential solutions.
Note that many of these issues are applicable across methods in ecology and evolution, not just PCMs.

@@

\subsubsection{Ornstein-Uhlenbeck (single stationary peak) models of trait evolution}
Most models of trait evolution are based on the Brownian motion model \citep{cavalli1967,felsenstein1973maximum}.
The Ornstein-Uhlenbeck (OU) model can be thought of as a modification of the Brownian model with an additional parameter that measures the strength of return towards a theoretical optimum shared across a clade or subset of species \citep{hansen1997stabilizing,Butler:2004aa}.
OU models have become increasingly popular as they tend to fit the data better than Brownian motion models, and have attractive biological interpretations \citep{cooper2016}.
For example, fit to an OU model has been seen as evidence of evolutionary constraints, stabilising selection, niche conservatism, and selective regimes \citep{Wiens:2010aa,beaulieu2012modeling,christin2013anatomical,mahler2013exceptional}.
However, the OU model has several well-known caveats \citep[see][]{ives2010phylogenetic,boettiger2012your,hansen2012interpreting,ho2013asymptotic,ho2014intrinsic}. 
For example, it is frequently incorrectly favoured over simpler models when using Likelihood ratio tests, particularly for small datasets that are commonly used in these analyses (the median number of taxa used for OU studies is 58; \citealp{cooper2016}). 
Additionally, very small amounts of error in datasets can result in an OU model being favoured over Brownian motion simply because OU can accommodate more variance towards the tips of the phylogeny, rather than due to any interesting biological process \citep{boettiger2012your,pennell2015model}.
Finally, the literature describing the OU model is clear that a simple explanation of clade-wide stabilising selection is unlikely to account for data fitting an OU model \citep[e.g.][]{hansen1997stabilizing,hansen2005assessing}, but users of the model often state that this is the case.
Unfortunately, these limitations are rarely taken into account in empirical studies.

\subsubsection{3. Trait-dependent diversification} 
Analyses of trait-dependent diversification are used to detect whether particular traits promote high rates of diversification, leading to some clades becoming more diverse than others \citep{nee1994reconstructed}. 
These kinds of analyses have been applied extensively in recent years to a variety of taxa and traits \citep[e.g.][]{Goldberg22102010,Price01052012,Givnish201455,ELE:ELE12168}.
Most use the binary-state speciation and extinction model (BiSSE) and related methods (the original BiSSE paper \citet{maddison2007estimating} has been cited 394 times; Google Scholar search 9th December 2015).
However, \citet{rabosky2015model} recently reevaluated the method and its assumptions and showed via simulations that a strong correlation between a trait and diversification rate can be inferred from a single diversification rate shift within a tree, even if the shift is unrelated to the trait of interest.
They suggest that many examples of trait-dependent diversification actually reflect this rate heterogeneity in trees, and thus are biologically meaningless.
Interestingly, these caveats are mentioned in earlier papers \citep{maddison2007estimating,fitzjohn2010quantitative,FitzJohn:2012aa} but were seemingly not widely understood given the shock reaction to \citet{rabosky2015model}.

@@

  \subsubsection{1. Not everything is mentioned in the literature}
    As scientists we mainly communicate our ideas through the literature. 
    Unfortunately, some of the information needed to properly apply PCMs is not found in the literature. 
    We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students, and within labs developing methods (and it is occasionally closer to fiction than fact!).
    Sometimes the folklore is based on best practice and includes tricks to get methods working, or useful rules-of-thumb; other times it is more opinion-based, but over time these opinions become rules. 
    Useful PCM folklore is often shared among developers, and among collaborating groups, but is not always shared outside of these circles. 
    When it is shared, it tends to be as email exchanges of ``dark advice'' that is not accessible to the rest of the community.

    One example of PCM folklore is that species with Studentized residuals $\pm 3$ are often omitted from regressions of phylogenetic independent contrasts, to avoid highly influential outliers affecting the results \citep[e.g.][]{GEB:GEB355}. 
    The rationale for this comes from \citet{jones1997optimum}, however, the $\pm 3$ cut-off is arbitrary, and barely mentioned in the original paper, but has become a rule-of-thumb for running these analyses (the paper has been cited \textgreater 100 times, mostly as a justification of this procedure).
    Another example is in the defaults of programs that perform PCMs. 
    These often start out as arbitrary starting points for data exploration with no justification for their use, but over time become the way the analysis is always performed.

    Other information about the limitations of a method may be absent from the literature due to the time-lag between a new method being published and others having time to test it. 
    For example, Felsenstein published the phylogenetic independent contrasts method in 1985 \citep{felsenstein1985phylogenies}, but it was not until the early 1990s that thorough critiques of the method and its assumptions began to be published \citep[e.g.][]{garland1992procedures}.
    This time-lag is shorter with more recent methods for a number of reasons.
    First, historically theory and software papers were generally separated, whereas currently they are combined in the same papers making it easier to run simulations. 
    Second, simulations testing methods are now required by journals (although simulations only show that the method behaves appropriately under ideal conditions), and finally, there are now far more people in the field and thus more papers published annually.
    However, even with this reduced time-lag, we suspect there are still many hidden assumptions and biases in all PCMs, even established methods, that have yet to be properly explored in the literature.
    For example, see Maddison \& FitzJohn's (\citeyear{maddison2014unsolved}) recent critique of Pagel's (\citeyear{pagel1994detecting}) correlated evolution method, and Rabosky \& Goldberg's (\citeyear{rabosky2015model}) discussion of trait-dependent speciation models \citep{nee1994reconstructed,maddison2007estimating}.

  \subsection{2. The literature is too technical and/or important details are difficult to locate}
    Although some information is not found in the literature (see above), the majority of assumptions and biases of PCMs are documented somewhere. 
    A big issue for novice methods-users (and often for advanced users too) is that this information can be extremely technical and dense.
    It is not unusual for papers to be long and full of equations. 
    Of course, such detail is critical for describing methods and facilitates testing/assessment, implementation and future developments of the method.
    Additionally, in most cases equations could not simply be replaced with text.
    There is evidence, however, that heavy use of equations impedes understanding and communication of concepts in biology \citep{fawcett2012heavy}.\\

    Another issue is that end-users need to understand the assumptions and caveats of methods.
    Within many comparative methods papers, assumptions and caveats can be found in the Introduction, Methods, Results and/or Discussion of a paper - they are rarely neatly corralled in one place.
    The sheer volume of literature can thus become a barrier to understanding, even, and perhaps especially, for the best known methods.
    For example, we recently reviewed papers discussing the assumptions and limitations of phylogenetic independent contrasts \citep{felsenstein1985phylogenies}, including how they are related to phylogenetic generalised least squares models \citep{garland2000using,rohle2006comment,blomberg2012independent}. 
    Even with prior knowledge of the key papers and authors to focus on, this resulted in $\approx 300$ manuscript pages and a book to read in order to fully understand the method and its caveats.
    The volume of reading itself is perhaps not the key issue, rather it is assessing when you have reached sufficient understanding which may not be clear to all users.\\

    The combined effects of a vast literature and sometimes opaque assumptions make it easy to miss pertinent details in PCM papers.
    These problems relate to due diligence for both end-users and developers. 
    Methods developers are not responsible for making sure that end-users read the literature. 
    Instead the onus is on the end-user to ensure that they have a clear understanding of the methods and caveats prior to using them. 
    However, simple steps could be taken by methods developers, such as subheadings that point to caveats and assumptions, to add clarity and limit method misuse and misinterpretation.

@@

    More recently however, the community has moved towards mostly implementing methods in R (\citealp{R-Core-Team:2014aa}; for a list of packages see Brian O'Meara's ``CRAN Task View: Phylogenetics, Especially Comparative Methods'' http://ftp.iitm.ac.in/cran/web/views/Phylogenetics.html), and code sharing has become almost ubiquitous. 
    The number of R packages for PCMs has increased markedly since 2005 when APE \citep{Paradis:2004aa} was released, and has increased particularly sharply since 2008 (Figure \ref{PCMRpackages}).
    Simultaneously, more people are able to use R thanks to workshops and changes in student training, thus when a new method is published it is now possible to take an R package ``off the shelf'' and use it to run the method immediately.
    The benefits of R are clear. 
    It is available to all, has a wide and engaged support community, and perhaps most importantly the source code for new methods is accessible.
    Users can fully explore any new method by examining the source code and running their own simulations. 
    Indeed, \citet{freckleton2009seven} suggested that the ability to conduct PCMs in flexible computing environments such as R would improve our ability to implement methods correctly. 
    However, the increasing use of R has instead led to more people (including the authors of this paper) jumping straight to the implementation of a method, without fully understanding what the method is doing, what its assumptions might be, or what the results mean in a biological context.
    %This is again an example of where users can and should conduct due diligence but also where developers and experienced users can help.
    This is not the fault of methods developers and users should conduct due diligence in understanding a method before using it. 
    Unfortunately the problem is exacerbated by the fact that manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
    The APE book \citep[an excellent resource;][]{paradis2011analysis} for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well-established and very easily implemented in R (for a counter example see \texttt{CAIC} and \texttt{caper} documentation; \citealp{purvis1995comparative,Orme:2013aa}).

\subsection{How can we solve some of these problems?}
  \subsubsection{1. Simplify, summarise and share}
    Many of the problems above arise because PCMs are difficult to communicate in purely descriptive terms.
    This is not entirely the fault of those writing these papers; most journals have strict word limits and a one line equation is generally a more succinct and precise descriptor of a method than a paragraph of text. 
    As we note above, the onus is on the end-user to read extensively and to do due diligence.
    Nonetheless, due diligence is a responsibility at all levels of research and that includes ensuring clarity to all target users, especially in describing assumptions and limitations.\\

@@

  \subsubsection{2. Factors to consider before using a method}
    A key skill to develop in science is cynicism, i.e., never take results from PCMs (or any other statistical analysis) at face value.
    As the Manic Street Preachers (\citeyear{manics}) put it, ``Cynicism is the only thing that keeps me sane''.
    At a minimum, users should read the original papers describing a method, plus any recent updates, and look carefully for assumptions and caveats that may affect the analyses at hand.
    A good way to check a method is to simulate some data and see if the results are as expected \citep[e.g.][]{boettiger2012your}. 
    This can expose hidden assumptions or biases that have not been explored in the papers accompanying the method, or reveal a lack of understanding of the the mechanics of the method being used. 
    It is also important to determine whether the method will work on a particular dataset.
    One key consideration is how many species are required for reasonable power. 
    Often methods require more species than are usually available. For example, the new trait-dependent diversification method of \citet{rabosky2015robust} is ``primarily applicable to phylogenies that include at least several thousand tips'' although the authors suggest that most empirical analyses have \textless 1000 tips. 
    Other considerations include whether the method is influenced by polytomies and whether the method is applicable to both ultrametric and non-ultrametric trees. 
    Indeed, many methods arbitrarily resolve polytomies using zero length branches, thus polytomies can inflate rates of evolution, and bias models of evolution  \citep{cooper2010body}. 
    Some current implementations of the OU model should not be used with non-ultrametric trees (e.g. MOTMOT; \citealp{Thomas:2011aa}) because they are based on transforming the tree directly, rather than transforming the variance co-variance matrix. 
    The problem is that where there is a pair of tips and at least one tip does not survive to the present, the expected covariances relating each of those two tips with any other tip in the tree are not identical. 
    Worked examples and explanations are provided in \citep{slater2014correction}. 
    Although this is not a problem with applying the OU model to non-ultrametric trees \textit{per se}, it is an example of different implementations of a common model that some users may not be aware of.

    It is also important to avoid retrofitting questions to the newest methods, instead we should think carefully about the question, whether the method is appropriate for the question, and whether PCMs are needed at all \citep{westoby1995misinterpreting,losos2011seeing}.
    In some cases, editors and reviewers may suggest using PCMs where they are not appropriate, and users should feel confident in rejecting these suggestions.
    Finally, end-users should never be afraid to question standard practice, sometimes it is just PCM folklore.

  \subsubsection{3. Solutions and incentives}
    It is important to recognise that our ability to do rigorous quantitative science often relies on highly-skilled methods-developers, especially as evolutionary biology becomes ever more computationally intensive. 
    We cannot afford to lose these people to industry, nor can we afford to pay industry wages, thus we need to make it worthwhile for such skilled researchers to remain in (or at least interact with) academia.

@@

    Finally, an obvious solution to many of these problems is for methods-developers and end-users to collaborate more. 
    Both parties can benefit greatly by collaboration. 
    Some benefits are obvious, e.g. methods-developers can gain extra datasets to test their ideas on and people who will discover corner cases and bugs in their software before it gets released, whereas end-users can work with the most cutting-edge methods and software.  
    Most methods are not designed in a vacuum, they have a specific purpose usually based on predictions from theory, experimentation or observation.
    There is a huge benefit in sharing ideas as well as products (data and code) and time, as long as the benefits to both parties are not heavily asymmetric.

>>>>

\parindent = 1.5em
\addtolength{\parskip}{.3em}
% Abstract 350 words max
\section{Abstract}
\begin{enumerate}
  \item Phylogenetic comparative methods are becoming increasingly popular for investigating evolutionary patterns and processes.
  However, these methods are not infallible - they suffer from biases and make assumptions like all other statistical methods.
  \item Unfortunately, although these limitations are generally well known in the phylogenetic comparative methods community, they are often inadequately assessed in empirical studies leading to misinterpreted results and poor model fits.
  Here we explore reasons for the communication gap dividing those developing new methods and those using them. 
  \item We suggest that some important pieces of information are missing from the literature, and that others are difficult to extract from long, technical papers.
  We also highlight problems with users jumping straight into software implementations of methods (e.g. in R) that may lack documentation on biases and assumptions that are mentioned in the original papers.
  \item To help solve these problems we make a number of suggestions including providing
  blog posts or videos to explain new methods in less technical terms, encouraging reproducibility and code sharing, making wiki-style pages summarising the literature on popular methods, more careful consideration and testing of whether a method is appropriate for a given question/dataset, increased collaboration, and a shift from publishing purely novel methods to publishing improvements to existing methods and ways of detecting biases or testing model fit.
  Many of these points are applicable across methods in ecology and evolution, not just phylogenetic comparative methods.
\end{enumerate}

@@

We suggest that one cause of this problem is that although researchers developing and implementing new methods are aware of the limitations of their methods and the assumptions that underly them, this information is not always being effectively transferred to end users.
Additionally, the tools and approaches used to fit models are often far more user-friendly and better documented than the methods used to assess whether that model fit is reasonable. 
Clearly more effort is needed to bridge the widening gap between those developing methods and end users. 
Here we explore the causes of this communication gap and suggest some potential solutions.
Note that many of these issues are applicable across methods in ecology and evolution, not just PCMs.

@@

\subsubsection{Ornstein-Uhlenbeck (single stationary peak) models of trait evolution}
Most models of trait evolution are based on the Brownian motion model \citep{cavalli1967,felsenstein1973maximum}.
The Ornstein-Uhlenbeck (OU) model can be thought of as a modification of the Brownian model with an additional parameter that measures the strength of return towards a theoretical optimum shared across a clade or subset of species \citep{hansen1997stabilizing,Butler:2004aa}.
OU models have become increasingly popular as they tend to fit the data better than Brownian motion models, and have attractive biological interpretations \citep{cooper2016}.
For example, fit to an OU model has been seen as evidence of evolutionary constraints, stabilising selection, niche conservatism, and selective regimes \citep{Wiens:2010aa,beaulieu2012modeling,christin2013anatomical,mahler2013exceptional}.
However, the OU model has several well known caveats \citep[see][]{ives2010phylogenetic,boettiger2012your,hansen2012interpreting,ho2013asymptotic,ho2014intrinsic}. 
For example, it is frequently incorrectly favoured over simpler models when using Likelihood ratio tests, particularly for small datasets that are commonly used in these analyses (the median number of taxa used for OU studies is 58; \citealp{cooper2016}). 
Additionally, very small amounts of error in datasets can result in an OU model being favoured over Brownian motion simply because OU can accommodate more variance towards the tips of the phylogeny, rather than due to any interesting biological process \citep{boettiger2012your,pennell2015model}.
Finally, the literature describing the OU model is clear that a simple explanation of clade-wide stabilising selection is unlikely to account for data fitting an OU model \citep[e.g.][]{hansen1997stabilizing,hansen2005assessing}, but users of the model often state that this is the case.
Unfortunately, these limitations are rarely taken into account in empirical studies.

\subsubsection{3. Trait-dependent diversification} 
Analyses of trait-dependent diversification are used to detect whether particular traits promote high rates of diversification, leading to some clades becoming more diverse than others \citep{nee1994reconstructed}. 
These kinds of analyses have been applied extensively in recent years to a variety of taxa and traits \citep[e.g.][]{Goldberg22102010,Price01052012,Givnish201455,ELE:ELE12168}.
Most use the binary state speciation and extinction model (BiSSE) and related methods (the original BiSSE paper \citet{maddison2007estimating} has been cited 394 times; Google Scholar search 9th December 2015).
However, \citet{rabosky2015model} recently reevaluated the method and its assumptions and showed via simulations that a strong correlation between a trait and diversification rate can be inferred from a single diversification rate shift within a tree, even if the shift is unrelated to the trait of interest.
They suggest that many examples of trait-dependent diversification actually reflect this rate heterogeneity in trees, and thus are biologically meaningless.
Interestingly, these caveats are mentioned in earlier papers \citep{maddison2007estimating,fitzjohn2010quantitative,FitzJohn:2012aa} but were seemingly not widely understood given the shock reaction to \citet{rabosky2015model}.

@@

  \subsubsection{1. Not everything is mentioned in the literature}
    As scientists we mainly communicate our ideas through the literature. 
    Unfortunately, some of the information needed to properly apply PCMs is not found in the literature. 
    We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students, and within labs developing methods (and it is occasionally closer to fiction than fact!).
    Sometimes the folklore is based on best practice and includes tricks to get methods working, or useful rules of thumb; other times it is more opinion based, but over time these opinions become rules. 
    Useful PCM folklore is often shared among developers, and among collaborating groups, but is not always shared outside of these circles. 
    When it is shared, it tends to be as email exchanges of ``dark advice'' that is not accessible to the rest of the community.

    One example of PCM folklore is that species with Studentized residuals $\pm 3$ are often omitted from regressions of phylogenetic independent contrasts, to avoid highly influential outliers affecting the results \citep[e.g.][]{GEB:GEB355}. 
    The rationale for this comes from \citet{jones1997optimum}, however, the $\pm 3$ cut-off is arbitrary, and barely mentioned in the original paper, but has become a rule of thumb for running these analyses (the paper has been cited \textgreater 100 times, mostly as a justification of this procedure).
    Another example is in the defaults of programs that perform PCMs. 
    These often start out as arbitrary starting points for data exploration with no justification for their use, but over time become the way the analysis is always performed.

    Other information about the limitations of a method may be absent from the literature due to the time lag between a new method being published and others having time to test it. 
    For example, Felsenstein published the phylogenetic independent contrasts method in 1985 \citep{felsenstein1985phylogenies}, but it was not until the early 1990s that thorough critiques of the method and its assumptions began to be published \citep[e.g.][]{garland1992procedures}.
    This time lag is shorter with more recent methods for a number of reasons.
    First, historically theory and software papers were generally separated, whereas currently they are combined in the same papers making it easier to run simulations. 
    Second, simulations testing methods are now required by journals (although simulations usually only show that the method behaves appropriately under ideal conditions), and finally, there are now far more people in the field and thus more papers published annually.
    However, even with this reduced time lag, we suspect there are still many hidden assumptions and biases in all PCMs, even established methods, that have yet to be properly explored in the literature.
    For example, see Maddison \& FitzJohn's (\citeyear{maddison2014unsolved}) recent critique of Pagel's (\citeyear{pagel1994detecting}) correlated evolution method, and Rabosky \& Goldberg's (\citeyear{rabosky2015model}) discussion of trait-dependent speciation models \citep{nee1994reconstructed,maddison2007estimating}.

  \subsection{2. The literature is too technical and/or important details are difficult to locate}
    Although some information is not found in the literature (see above), the majority of assumptions and biases of PCMs are documented somewhere. 
    A big issue for novice methods users (and often for advanced users too) is that this information can be extremely technical and dense.
    It is not unusual for papers to be long and full of equations. 
    Of course, such detail is critical for describing methods and facilitates testing/assessment, implementation and future developments of the method.
    Additionally, in most cases equations could not simply be replaced with text.
    There is evidence, however, that heavy use of equations impedes understanding and communication of concepts in biology \citep{fawcett2012heavy}.\\

    Another issue is that end users need to understand the assumptions and caveats of methods.
    Within many comparative methods papers, assumptions and caveats can be found in the Introduction, Methods, Results and/or Discussion of a paper - they are rarely neatly corralled in one place.
    The sheer volume of literature can thus become a barrier to understanding, even, and perhaps especially, for the best known methods.
    For example, we recently reviewed papers discussing the assumptions and limitations of phylogenetic independent contrasts \citep{felsenstein1985phylogenies}, including how they are related to phylogenetic generalised least squares models \citep{garland2000using,rohle2006comment,blomberg2012independent}. 
    Even with prior knowledge of the key papers and authors to focus on, this resulted in $\approx 300$ manuscript pages and a book to read in order to fully understand the method and its caveats.
    The volume of reading itself is perhaps not the key issue, rather it is assessing when you have reached sufficient understanding which may not be clear to all users.\\

    The combined effects of a vast literature and sometimes opaque assumptions make it easy to miss pertinent details in PCM papers.
    These problems relate to due diligence for both end users and developers. 
    Methods developers are not responsible for making sure that end users read the literature. 
    Instead the onus is on the end user to ensure that they have a clear understanding of the methods and caveats prior to using them. 
    However, simple steps could be taken by methods developers, such as subheadings that point to caveats and assumptions, to add clarity and limit method misuse and misinterpretation.

@@

    More recently however, the community has moved towards mostly implementing methods in R (\citealp{R-Core-Team:2014aa}; for a list of packages see Brian O'Meara's ``CRAN Task View: Phylogenetics, Especially Comparative Methods'' https://cran.r-project.org/web/views/Phylogenetics.html), and code sharing has become almost ubiquitous. 
    The number of R packages for PCMs has increased markedly since 2005 when APE \citep{Paradis:2004aa} was released, and has increased particularly sharply since 2008 (Figure \ref{PCMRpackages}).
    Simultaneously, more people are able to use R thanks to workshops and changes in student training, thus when a new method is published it is now possible to take an R package ``off the shelf'' and use it to run the method immediately.
    The benefits of R are clear. 
    It is available to all, has a wide and engaged support community, and perhaps most importantly the source code for new methods is accessible.
    Users can fully explore any new method by examining the source code and running their own simulations. 
    Indeed, \citet{freckleton2009seven} suggested that the ability to conduct PCMs in flexible computing environments such as R would improve our ability to implement methods correctly. 
    However, the increasing use of R has instead led to more people (including the authors of this paper) jumping straight to the implementation of a method, without fully understanding what the method is doing, what its assumptions might be, or what the results mean in a biological context.
    %This is again an example of where users can and should conduct due diligence but also where developers and experienced users can help.
    This is not the fault of methods developers and users should conduct due diligence in understanding a method before using it. 
    Unfortunately the problem is exacerbated by the fact that manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
    The APE book \citep[an excellent resource;][]{paradis2011analysis} for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well established and very easily implemented in R (for a counter example see \texttt{CAIC} and \texttt{caper} documentation; \citealp{purvis1995comparative,Orme:2013aa}).

\subsection{How can we solve some of these problems?}
  \subsubsection{1. Simplify, summarise and share}
    Many of the problems above arise because PCMs are difficult to communicate in purely descriptive terms.
    This is not entirely the fault of those writing these papers; most journals have strict word limits and a one line equation is generally a more succinct and precise descriptor of a method than a paragraph of text. 
    As we note above, the onus is on the end user to read extensively and to do due diligence.
    Nonetheless, due diligence is a responsibility at all levels of research and that includes ensuring clarity to all target users, especially in describing assumptions and limitations.\\

@@

  \subsubsection{2. Factors to consider before using a method}
    A key skill to develop in science is cynicism, i.e., never take results from PCMs (or any other statistical analysis) at face value.
    As the Manic Street Preachers (\citeyear{manics}) put it, ``Cynicism is the only thing that keeps me sane''.
    At a minimum, users should read the original papers describing a method, plus any recent updates, and look carefully for assumptions and caveats that may affect the analyses at hand.
    A good way to check a method is to simulate some data and see if the results are as expected \citep[e.g.][]{boettiger2012your}. 
    This can expose hidden assumptions or biases that have not been explored in the papers accompanying the method, or reveal a lack of understanding of the the mechanics of the method being used. 
    It is also important to determine whether the method will work on a particular dataset.
    One key consideration is how many species are required for reasonable power. 
    Often methods require more species than are usually available. For example, the new trait-dependent diversification method of \citet{rabosky2015robust} is ``primarily applicable to phylogenies that include at least several thousand tips'' although the authors suggest that most empirical analyses have \textless 1000 tips. 
    Other considerations include whether the method is influenced by polytomies and whether the method is applicable to both ultrametric and non-ultrametric trees. 
    Indeed, many methods arbitrarily resolve polytomies using zero length branches, thus polytomies can inflate rates of evolution, and bias models of evolution  \citep{cooper2010body}. 
    Some current implementations of the OU model should not be used with non-ultrametric trees (e.g. MOTMOT; \citealp{Thomas:2011aa}) because they are based on transforming the tree directly, rather than transforming the variance covariance matrix. 
    The problem is that where there is a pair of tips and at least one tip does not survive to the present, the expected covariances relating each of those two tips with any other tip in the tree are not identical. 
    Worked examples and explanations are provided in \citep{slater2014correction}. 
    Although this is not a problem with applying the OU model to non-ultrametric trees \textit{per se}, it is an example of different implementations of a common model that some users may not be aware of.

    It is also important to avoid retrofitting questions to the newest methods, instead we should think carefully about the question, whether the method is appropriate for the question, and whether PCMs are needed at all \citep{westoby1995misinterpreting,losos2011seeing}.
    In some cases, editors and reviewers may suggest using PCMs where they are not appropriate, and users should feel confident in rejecting these suggestions.
    Finally, end users should never be afraid to question standard practice, sometimes it is just PCM folklore.

  \subsubsection{3. Solutions and incentives}
    It is important to recognise that our ability to do rigorous quantitative science often relies on highly-skilled methods developers, especially as evolutionary biology becomes ever more computationally intensive. 
    We cannot afford to lose these people to industry, nor can we afford to pay industry wages, thus we need to make it worthwhile for such skilled researchers to remain in (or at least interact with) academia.

@@

    Finally, an obvious solution to many of these problems is for methods developers and end users to collaborate more. 
    Both parties can benefit greatly by collaboration. 
    Some benefits are obvious, e.g. methods developers can gain extra datasets to test their ideas on and people who will discover corner cases and bugs in their software before it gets released, whereas end users can work with the most cutting edge methods and software.  
    Most methods are not designed in a vacuum, they have a specific purpose usually based on predictions from theory, experimentation or observation.
    There is a huge benefit in sharing ideas as well as products (data and code) and time, as long as the benefits to both parties are not heavily asymmetric.
